---
title: "Hands-on Exercise 1"
author: "Huynh Minh Phuong"
---

# Getting Started

## Install and launch R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages have been installed. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

------------------------------------------------------------------------

## Importing the data

```{r}
exam_data <-read_csv("data/Exam_data.csv")
```

# Plotting

## Plotting a simple bar chart

```{r}
ggplot(data=exam_data,
       aes(y=forcats::fct_infreq(RACE)))+
  geom_bar()+
  theme_minimal()+
  theme(
    panel.background = element_rect(fill = "lightblue", linetype=0))+
  labs(y="Race",
       x="Count")
  
```

## Sorted bar charts with frequencies and percentages

```{r}
ggplot(data=exam_data,
       aes(x=forcats::fct_infreq(RACE)))+
  geom_bar()+
  ylim(0,220)+
  labs(x="Race",
       y="No. of\nPupils")+
  theme(axis.title.y=element_text(angle=0))+
  geom_text(stat="count", 
      aes(label=paste0(..count.., ", ", 
      round(..count../sum(..count..)*100, 1), "%")),
      vjust=-1)
```

## Adding mean and median lines on histogram plot

```{r}
ggplot(data=exam_data,
       aes(x=MATHS))+
  geom_histogram(bins=20, 
                 color='black', 
                 fill='lightblue')+
    labs(y="Count",
       x="Maths")+
    geom_vline(aes(xintercept=mean(MATHS, na.rm=T)),
             color="red", 
             linetype="dashed", 
             size=1)+
    geom_vline(aes(xintercept=median(MATHS, na.rm=T)),
             color="grey30",
             linetype="dashed", 
             size=1)
  
```

## Background histogram for added context

```{r, warning=FALSE, message=FALSE}
ggplot(data=exam_data,
       aes(x=ENGLISH, fill=GENDER))+
  geom_histogram(data=exam_data[,-3], fill='grey', alpha=0.5)+
  geom_histogram(colour='black')+
  facet_wrap(~GENDER)+
  guides(fill="none")+
  theme_bw()
```

## Relationship between English and Maths scores

```{r}
ggplot(data=exam_data,
       aes(x=MATHS, y=ENGLISH))+
  geom_point()+
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+
  geom_vline(aes(xintercept=50),
             color="grey60", 
             linetype="dashed", 
             size=1)+
  geom_hline(yintercept=50,
             linetype="dashed",
             color="grey60",
             size=1)
```
