---
title: "Hands-on Exercise 6"
author: "Huynh Minh Phuong"
---

# Creating Ternary Plot

Ternary plots are a way of displaying the distribution and variability of three-part compositional data. (For example, the proportion of aged, economy active and young population or sand, silt, and clay in soil.) It's display is a triangle with sides scaled from 0 to 1. Each side represents one of the three components. A point is plotted so that a line drawn perpendicular from the point to each leg of the triangle intersect at the component values of the point.

## Getting started

**ggtern is** a ggplot extension specially designed to plot ternary diagrams. The package will be used to plot static ternary plots.

```{r}
pacman::p_load(plotly, ggstatsplot,tidyverse, ggtern)
```

Import data

```{r}
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv")
```

## Prepare the data

Use the ***mutate()*** function of **dplyr** package to derive three new measures, namely: young, active, and old.

```{r}
#Deriving the young, economy active and old measures
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year))%>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8]))%>%
  mutate(ACTIVE = rowSums(.[9:16]))  %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018)%>%
  filter(TOTAL > 0)

```

## **Plotting a static ternary diagram**

### Simple plot

```{r}
#Building the static ternary plot
ggtern(data=agpop_mutated,aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()
```

### Static ternary plot

```{r}
#Building the static ternary plot
ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()
```

### Interactive ternary diagram using plot_ly()

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  )
```

# Visual Correlation Analysis

Correlation coefficient is a popular statistic that use to measure the type and strength of the relationship between two variables. The values of a correlation coefficient ranges between -1.0 and 1.0. A correlation coefficient of 1 shows a perfect linear relationship between the two variables, while a -1.0 shows a perfect inverse relationship between the two variables. A correlation coefficient of 0.0 shows no linear relationship between the two variables.

When multivariate data are used, the correlation coefficeints of the pair comparisons are displayed in a table form known as correlation matrix or scatterplot matrix.

There are three broad reasons for computing a correlation matrix.

-   To reveal the relationship between high-dimensional variables pair-wisely.

-   To input into other analyses. For example, people commonly use correlation matrices as inputs for exploratory factor analysis, confirmatory factor analysis, structural equation models, and linear regression when excluding missing values pairwise.

-   As a diagnostic when checking other analyses. For example, with linear regression a high amount of correlations suggests that the linear regression's estimates will be unreliable.

When the data is large, both in terms of the number of observations and the number of variables, [Corrgram](http://www.datavis.ca/papers/corrgram.pdf) tend to be used to visually explore and analyse the structure and the patterns of relations among variables. It is designed based on two main schemes:

-   Rendering the value of a correlation to depict its sign and magnitude, and

-   Reordering the variables in a correlation matrix so that "similar" variables are positioned adjacently, facilitating perception.

## Getting started

Load packages

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

```{r}
detach(package:ggtern, unload = TRUE)
```

Import data

```{r}
wine <- read_csv("data/wine_quality.csv")
glimpse(wine)
```

## Build Correlation Matrix: *pairs()* method

### Basic correlation matrix

```{r}
pairs(wine[,1:11])
```

The required input of *pairs()* can be a matrix or data frame. The code chunk used to create the scatterplot matrix is relatively simple. It uses the default *pairs* function. Columns 2 to 12 of wine dataframe is used to build the scatterplot matrix. The variables are: fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates and alcohol.

```{r}
pairs(wine[,2:12])
```

### Draw lower corner

It is a common practice to show either the upper half or lower half of the correlation matrix instead of both. This is because a correlation matrix is symmetric. To show the lower half of the correlation matrix, the upper.panel argument will be used as shown in the code chunk below.

```{r}
pairs(wine[,2:12], upper.panel = NULL)
```

```{r}
pairs(wine[,2:12], lower.panel = NULL)
```

### Include correlation coefficients

To show the correlation coefficient of each pair of variables instead of a scatter plot, panel.cor function will be used. This will also show higher correlations in a larger font.

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)

```

## Visualize correlation matrix with ggcormat()

The sample sub-code chunk can be used to control specific component of the plot such as the font size of the x-axis, y-axis, and the statistical report.

```{r}
ggplot.component = list(
    theme(text=element_text(size=5),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8)))
```

### Multiple plots

```{r}
#| fig-height: 20
#| fig-width: 40
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

Things to learn from the code chunk above:

-   `cor.vars` argument is used to compute the correlation matrix needed to build the corrgram.

-   `ggcorrplot.args` argument provide additional (mostly aesthetic) arguments that will be passed to [`ggcorrplot::ggcorrplot`](http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-matrix-using-ggplot2) function. The list should avoid any of the following arguments since they are already internally being used: `corr`, `method`, `p.mat`, `sig.level`, `ggtheme`, `colors`, `lab`, `pch`, `legend.title`, `digits`.

-   to build a facet plot, the only argument needed is `grouping.var`.

-   Behind *group_ggcorrmat()*, **patchwork** package is used to create the multiplot. `plotgrid.args` argument provides a list of additional arguments passed to [*patchwork::wrap_plots*](https://patchwork.data-imaginist.com/reference/wrap_plots.html), except for guides argument which is already separately specified earlier.

## **Visualising Correlation Matrix using corrplot Package**

### **Getting started with corrplot**

[*cor()*](https://www.rdocumentation.org/packages/stats/versions/3.6.0/topics/cor) of R Stats is used to compute the correlation matrix of wine data frame.

```{r}
wine.cor <- cor(wine[, 1:11])
```

[*corrplot()*](https://www.rdocumentation.org/packages/corrplot/versions/0.2-0/topics/corrplot) is used to plot the corrgram by using all the default setting as shown in the code chunk below

```{r}
corrplot(wine.cor)
```

### **Working with visual geometrics**

In **corrplot** package, there are seven visual geometrics (parameter method) can be used to encode the attribute values. They are: circle, square, ellipse, number, shade, color and pie. The default is circle. As shown in the previous section, the default visual geometric of corrplot matrix is circle. However, this default setting can be changed by using the *method* argument as shown in the code chunk below.

```{r}
corrplot(wine.cor, 
         method = "ellipse") 
```

*corrplor()* supports three layout types, namely: \"full\", \"upper\" or \"lower\". The default is \"full\" which display full matrix. The default setting can be changed by using the *type* argument of *corrplot()*.

### **Working with layout**

*corrplor()* supports three layout types, namely: \"full\", \"upper\" or \"lower\". The default is \"full\" which display full matrix. The default setting can be changed by using the *type* argument of *corrplot()*.

The default layout of the corrgram can be further customised. For example, arguments *diag* and *tl.col* are used to turn off the diagonal cells and to change the axis text label colour to black colour respectively as shown in the code chunk and figure below.

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower",
         diag = FALSE,
         tl.col = "black")
```

### **Working with mixed layout**

With **corrplot** package, it is possible to design corrgram with mixed visual matrix of one half and numerical matrix on the other half. In order to create a coorgram with mixed layout, the corrplot.mixed(), a wrapped function for mixed visualisation style will be used.

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

### **Combining corrgram with the significant test**

With corrplot package, we can use the *cor.mtest()* to compute the p-values and confidence interval for each pair of variables.

```{r}
wine.sig = cor.mtest(wine.cor, conf.level= .95)
```

We can then use the *p.mat* argument of *corrplot* function as shown in the code chunk below.

```{r}
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```

### Reorder a corrgram

Matrix reorder is very important for mining the hiden structure and pattern in a corrgram. By default, the order of attributes of a corrgram is sorted according to the correlation matrix (i.e. \"original\"). The default setting can be over-write by using the *order* argument of *corrplot()*. Currently, **corrplot** package support four sorting methods, they are:

-   \"AOE\" is for the angular order of the eigenvectors. See Michael Friendly (2002) for details.

-   \"FPC\" for the first principal component order.

-   \"hclust\" for hierarchical clustering order, and \"hclust.method\" for the agglomeration method to be used.

    -   \"hclust.method\" should be one of \"ward\", \"single\", \"complete\", \"average\", \"mcquitty\", \"median\" or \"centroid\".

-   \"alphabet\" for alphabetical order.

\"AOE\", \"FPC\", \"hclust\", \"alphabet\". More algorithms can be found in **seriation** package.

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")
```

### **Reordering a correlation matrix using hclust**

If using **hclust**, ***corrplot()*** can draw rectangles around the corrgram based on the results of hierarchical clustering.

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```

# Heatmap for Visualising and Analysing Multivariate Data

## Getting Started

Load packages:

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

## Import data set

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

Prepare the data

Next, we need to change the rows by country name instead of row number by using the code chunk below

```{r}
row.names(wh) <- wh$Country
```

The data was loaded into a data frame, but it has to be a data matrix to make your heatmap.

```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

## Static Heatmap

There are many R packages and functions can be used to drawing static heatmaps, they are:

-   [heatmap()](https://www.rdocumentation.org/packages/stats/versions/3.6.0/topics/heatmap)of R stats package. It draws a simple heatmap.

-   [heatmap.2()](https://www.rdocumentation.org/packages/gplots/versions/3.0.1.1/topics/heatmap.2) of **gplots** R package. It draws an enhanced heatmap compared to the R base function.

-   [pheatmap()](https://www.rdocumentation.org/packages/pheatmap/versions/1.0.12/topics/pheatmap) of [**pheatmap**](https://www.rdocumentation.org/packages/pheatmap/versions/1.0.12) R package. **pheatmap** package also known as Pretty Heatmap. The package provides functions to draws pretty heatmaps and provides more control to change the appearance of heatmaps.

-   [**ComplexHeatmap**](https://bioconductor.org/packages/release/bioc/html/ComplexHeatmap.html) package of R/Bioconductor package. The package draws, annotates and arranges complex heatmaps (very useful for genomic data analysis). The full reference guide of the package is available [here](https://jokergoo.github.io/ComplexHeatmap-reference/book/).

-   [**superheat**](https://cran.r-project.org/web/packages/superheat/) package: A Graphical Tool for Exploring Complex Datasets Using Heatmaps. A system for generating extendable and customizable heatmaps for exploring complex datasets, including big data and data with multiple data types. The full reference guide of the package is available [here](https://rlbarter.github.io/superheat/).

### heatmap() of R Stats

By default, **heatmap()** plots a cluster heatmap. The arguments ***Rowv=NA*** and ***Colv=NA*** are used to switch off the option of plotting the row and column dendrograms.

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)
```

::: callout-note
The order of both rows and columns is different compare to the native wh_matrix. This is because heatmap do a reordering using clusterisation: it calculates the distance between each pair of rows and columns and try to order them by similarity. Moreover, the corresponding dendrogram are provided beside the heatmap.
:::

Here, red cells denotes small values, and red small ones. This heatmap is not really informative. Indeed, the Happiness Score variable have relatively higher values, what makes that the other variables with small values all look the same. Thus, we need to normalize this matrix. This is done using the *scale* argument. It can be applied to rows or to columns following your needs.

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

## Interactive Heatmap

### **Working with heatmaply**

-   Different from *heatmap()*, for *heatmaply()* the default horizontal dendrogram is placed on the left hand side of the heatmap.

-   The text label of each raw, on the other hand, is placed on the right hand side of the heat map.

-   When the x-axis marker labels are too long, they will be rotated by 135 degree from the north.

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```

### Data transformation

### Scaling method

-   When all variables are came from or assumed to come from some normal distribution, then scaling (i.e.: subtract the mean and divide by the standard deviation) would bring them all close to the standard normal distribution.

-   In such a case, each value would reflect the distance from the mean in units of standard deviation.

-   The *scale* argument in *heatmaply()* supports column and row scaling.

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column")
```

### Normalizing method

-   When variables in the data comes from possibly different (and non-normal) distributions, the normalize function can be used to bring data to the 0 to 1 scale by subtracting the minimum and dividing by the maximum of all observations.

-   This preserves the shape of each variable\'s distribution while making them easily comparable on the same \"scale\".

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
```

### Percentising method

-   This is similar to ranking the variables, but instead of keeping the rank values, divide them by the maximal rank.

-   This is done by using the ecdf of the variables on their own values, bringing each value to its empirical percentile.

-   The benefit of the percentize function is that each value has a relatively clear interpretation, it is the percent of observations that got that value or below it.

```{r}
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```

### Clustering algorithm

**heatmaply** supports a variety of hierarchical clustering algorithm. The main arguments provided are:

-   *distfun*: function used to compute the distance (dissimilarity) between both rows and columns. Defaults to dist. The options \"pearson\", \"spearman\" and \"kendall\" can be used to use correlation-based clustering, which uses as.dist(1 - cor(t(x))) as the distance metric (using the specified correlation method).

-   *hclustfun*: function used to compute the hierarchical clustering when *Rowv* or *Colv* are not dendrograms. Defaults to *hclust*.

-   *dist_method* default is NULL, which results in \"euclidean\" to be used. It can accept alternative character strings indicating the method to be passed to distfun. By default *distfun* is \"dist\"\" hence this can be one of \"euclidean\", \"maximum\", \"manhattan\", \"canberra\", \"binary\" or \"minkowski\".

-   *hclust_method* default is NULL, which results in \"complete\" method to be used. It can accept alternative character strings indicating the method to be passed to *hclustfun*. By default hclustfun is hclust hence this can be one of \"ward.D\", \"ward.D2\", \"single\", \"complete\", \"average\" (= UPGMA), \"mcquitty\" (= WPGMA), \"median\" (= WPGMC) or \"centroid\" (= UPGMC).

In general, a clustering model can be calibrated either manually or statistically.

### Manual approach

In the code chunk below, the heatmap is plotted by using hierachical clustering algorithm with \"Euclidean distance\" and \"ward.D\" method.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```

### Statistical approach

In order to determine the best clustering method and number of cluster the *dend_expend()* and *find_k()* functions of **dendextend** package will be used.

First, the *dend_expend()* will be used to determine the recommended clustering method to be used.

```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

Next, *find_k()* is used to determine the optimal number of cluster.

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

Figure above shows that k=3 would be good.

With reference to the statistical analysis results, we can prepare the code chunk as shown below

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```

### Seriation

One of the problems with hierarchical clustering is that it doesn\'t actually place the rows in a definite order, it merely constrains the space of possible orderings. Take three items A, B and C. If you ignore reflections, there are three possible orderings: ABC, ACB, BAC. If clustering them gives you ((A+B)+C) as a tree, you know that C can\'t end up between A and B, but it doesn\'t tell you which way to flip the A+B cluster. It doesn\'t tell you if the ABC ordering will lead to a clearer-looking heatmap than the BAC ordering.

**heatmaply** uses the seriation package to find an optimal ordering of rows and columns. Optimal means to optimize the Hamiltonian path length that is restricted by the dendrogram structure. This, in other words, means to rotate the branches so that the sum of distances between each adjacent leaf (label) will be minimized. This is related to a restricted version of the travelling salesman problem.

Here we meet our first seriation algorithm: Optimal Leaf Ordering (OLO). This algorithm starts with the output of an agglomerative clustering algorithm and produces a unique ordering, one that flips the various branches of the dendrogram around so as to minimize the sum of dissimilarities between adjacent leaves. Here is the result of applying Optimal Leaf Ordering to the same clustering result as the heatmap above.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

The default options is \"OLO\" (Optimal leaf ordering) which optimizes the above criterion (in O(n\^4)). Another option is \"GW\" (Gruvaeus and Wainer) which aims for the same goal but uses a potentially faster heuristic.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```

### Working with colour palletes

The default colour palette uses by **heatmaply** is *viridis*. heatmaply users, however, can use other colour palettes in order to improve the aestheticness and visual friendliness of the heatmap.

In the code chunk below, the Blues colour palette of rColorBrewer is used

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```

### Finishing touch

Beside providing a wide collection of arguments for meeting the statistical analysis needs, *heatmaply* also provides many plotting features to ensure cartographic quality heatmap can be produced.

In the code chunk below the following arguments are used:

-   *k_row* is used to produce 5 groups.

-   *margins* is used to change the top margin to 60 and row margin to 200.

-   *fontsizw_row* and *fontsize_col* are used to change the font size for row and column labels to 4.

-   *main* is used to write the main title of the plot.

-   *xlab* and *ylab* are used to write the x-axis and y-axis labels respectively.

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```

# **Visual Multivariate Analysis with Parallel Coordinates Plot**

## Getting Started

Load packages

For this exercise, the **GGally**, **parcoords**, **parallelPlot** and **tidyverse** packages will be used.

The code chunks below are used to install and load the packages in R.

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

Load data

In this hands-on exercise, the World Happinees 2018 (http://worldhappiness.report/ed/2018/) data will be used. The data set is download at https://s3.amazonaws.com/happiness-report/2018/WHR2018Chapter2OnlineData.xls. The original data set is in Microsoft Excel format. It has been extracted and saved in csv file called **WHData-2018.csv**.

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

## **Plotting Static Parallel Coordinates Plot**

Code chunk below shows a typical syntax used to plot a basic static parallel coordinates plot by using `ggparcoord()`.

```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

### **Plotting a parallel coordinates with boxplot**

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

-   `groupColumn` argument is used to group the observations (i.e. parallel lines) by using a single variable (i.e. Region) and colour the parallel coordinates lines by region name.

-   `scale` argument is used to scale the variables in the parallel coordinate plot by using `uniminmax` method. The method univariately scale each variable so the minimum of the variable is zero and the maximum is one.

-   `alphaLines` argument is used to reduce the intensity of the line colour to 0.2. The permissible value range is between 0 to 1.

-   `boxplot` argument is used to turn on the boxplot by using logical `TRUE`. The default is `FALSE`.

-   `title` argument is used to provide the parallel coordinates plot a title.

### **Parallel coordinates with facet**

In the code chunk below, `facet_wrap()`of **ggplot2** is used to plot 10 small multiple parallel coordinates plots. Each plot represent one geographical region such as East Asia.

To make the x-axis text label easy to read, let us rotate the labels by 30 degrees. We can rotate axis text labels using `theme()` function in ggplot2 as shown in the code chunk below

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

## Interactive Parallel Coordiantes Plot with parallelPlot

### The basic plot

```{r}
wh <- wh %>%
  select("Happiness score", c(7:12))
```

-   One of the useful interactive feature of parallelPlot is we can click on a variable of interest, for example Happiness score, the monotonous blue colour (default) will change a blues with different intensity colour scheme will be used.

-   We can change the default blue colour scheme by using `continousCS` argument.

-   `rotateTitle` argument is used to avoid overlapping axis labels.

```{r}
parallelPlot(wh,
             width = 320,
             height = 250,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

### **Parallel coordinates plot with histogram**

```{r}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```

# **Treemap Visualisation**

## Getting started

Load packages

```{r}
pacman::p_load(treemap, treemapify, tidyverse) 
```

Import data set

In this exercise, *REALIS2018.csv* data will be used. This dataset provides information of private property transaction records in 2018. The dataset is extracted from REALIS portal (https://spring.ura.gov.sg/lad/ore/login/index.cfm) of Urban Redevelopment Authority (URA).

```{r}
realis2018 <- read_csv("data/realis2018.csv")
```

## Data wrangling and manipulation

### **Grouped summaries without the Pipe**

Aggregation functions such as *sum()* and *meadian()* obey the usual rule of missing values: if there\'s any missing value in the input, the output will be a missing value. The argument *na.rm = TRUE* removes the missing values prior to computation.

```{r}
realis2018_grouped <- group_by(realis2018, `Project Name`,`Planning Region`, `Planning Area`,               `Property Type`, `Type of Sale`)

realis2018_summarised <- summarise(realis2018_grouped, 
                          `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
                          `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
                          `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE), 
                          `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

The code chunk above is not very efficient because we have to give each intermediate data.frame a name, even though we don\'t have to care about it.

### **Grouped summaries with the *pipe***

```{r}
realis2018_summarised <- realis2018 %>% 
  group_by(`Project Name`,`Planning Region`, 
           `Planning Area`, `Property Type`, 
           `Type of Sale`) %>%
  summarise(`Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE), 
            `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
            `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
            `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

## Design Treemap with treemap 

In this section, *treemap()* of **Treemap** package is used to plot a treemap showing the distribution of median unit prices and total unit sold of resale condominium by geographic hierarchy in 2017.

First, we will select records belongs to resale condominium property type from *realis2018_selected* data frame.

```{r}
realis2018_selected <- realis2018_summarised %>%
  filter(`Property Type` == "Condominium", `Type of Sale` == "Resale")
```

### The " value" type treemap

-   index

    -   The index vector must consist of at least two column names or else no hierarchy treemap will be plotted.

    -   If multiple column names are provided, such as the code chunk above, the first name is the highest aggregation level, the second name the second highest aggregation level, and so on.

-   vSize

    -   The column must not contain negative values. This is because it\'s vaues will be used to map the sizes of the rectangles of the treemaps.

-   The rectangles are coloured with different intensity of green, reflecting their respective median unit prices.

-   The legend reveals that the values are binned into ten bins, i.e. 0-5000, 5000-10000, etc. with an equal interval of 5000.

-   There are two arguments that determine the mapping to color palettes: *mapping* and *palette*. The only difference between \"value\" and \"manual\" is the default value for mapping. The \"value\" treemap considers palette to be a diverging color palette (say ColorBrewer\'s \"RdYlBu\"), and maps it in such a way that 0 corresponds to the middle color (typically white or yellow), -max(abs(values)) to the left-end color, and max(abs(values)), to the right-end color. The \"manual\" treemap simply maps min(values) to the left-end color, max(values) to the right-end color, and mean(range(values)) to the middle color.

-   although the colour palette used is RdYlBu but there are no red rectangles in the treemap above. This is because all the median unit prices are positive.

-   The reason why we see only 5000 to 45000 in the legend is because the *range* argument is by default c(min(values, max(values)) with some pretty rounding.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

### The "manual" type treemap

The \"manual\" type does not interpret the values as the \"value\" type does. Instead, the value range is mapped linearly to the colour palette.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

### **Working with *algorithm* argument**

When \"pivotSize\" algorithm is used, *sortID* argument can be used to dertemine the order in which the rectangles are placed from top left to bottom right.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "pivotSize",
        sortID = "Median Transacted Price",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## Design Treemap with treemapify

**treemapify** is a R package specially developed to draw treemaps in **ggplot2**. In this section, you will learn how to designing treemps closely resemble treemaps designing in previous section by using **treemapify**. Before you getting started, you should read [Introduction to \"treemapify\"](https://cran.r-project.org/web/packages/treemapify/vignettes/introduction-to-treemapify.html) its [user guide](https://cran.r-project.org/web/packages/treemapify/treemapify.pdf).

### Basic treemap

Group by Planning Region

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`),
       start = "topleft") + 
  geom_treemap()
```

Group by Planning Area

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap() +
  geom_treemap_subgroup2_border(colour = "gray40",
                                size = 2) +
  geom_treemap_subgroup_border(colour = "gray20")
```

## Design Treemap with d3treeR 

```{r}
library(devtools)
install_github("timelyportfolio/d3treeR")
```

```{r}
library(d3treeR)
```

1.  *treemap()* is used to build a treemap by using selected variables in condominium data.frame. The treemap created is save as object called ***tm***.

```{r}
tm <- treemap(realis2018_summarised,
        index=c("Planning Region", "Planning Area"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        title="Private Residential Property Sold, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

2.  Then *d3tree()* is used to build an interactive treemap.

```{r}
d3tree(tm,rootname = "Singapore" )
```
