---
title: "Hands-on Exercise 2"
author: "Huynh Minh Phuong"
---

# Getting Started

## Installing and loading the required libraries

Code chunk below will be used to check if these packages have been installed and also will load them onto your working R environment.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse)
```

------------------------------------------------------------------------

## Importing the data

```{r, warning=FALSE, message=FALSE}
exam_data <-read_csv("data/Exam_data.csv")
```

# Beyond ggplot 2 Fundamentals

## Anotation

### Anotation with ggrepel

Replace geom_text() by geom_text_repel() and geom_label() by geom_label_repel.

```{r, warning=FALSE, message=FALSE}
ggplot(data=exam_data,
       aes(y=ENGLISH,x=MATHS))+
  geom_point()+
  ggtitle("English scores versus Maths scores for Primary 3")+
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))+
  geom_smooth(method=lm, 
              size=0.5)+
  geom_label_repel(aes(label = ID), 
                   fontface = "bold")
  
```

## Theme

### Working with ggtheme package

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

### Working with hrbthems package

```{r, warning=FALSE}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
  
```

## Composite graphs

### Pathwork methods

#### Creating a composite figure with tag

```{r, warning=FALSE, message=FALSE}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")

p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")

p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")


#“|” operator to stack two ggplot2 graphs
#“/” operator to place the plots beside each other
#“()” operator the define the sequence of the plotting
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

#### Creating figure with insert

```{r, warning=FALSE, message=FALSE}
p12 <- p1|p2
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

#### Creating a composite figure by using patchwork and ggtheme

```{r, warning=FALSE, message=FALSE}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
